---
title: "Diabetes Risk Prediction Using Logistic Regression"
author: "Alwaleed Al Masqari"
date: "`r Sys.Date()`"
output: html_document
---

## Introduction
This project aims to predict the risk of diabetes using logistic regression based on medical and lifestyle factors. The analysis uses the Pima Indians Diabetes dataset.

## Load Libraries
```{r}
# install.packages(c("tidyverse", "caret"))
library(tidyverse)
library(caret)
```

## Load Data
```{r}
data <- read.csv("diabetes.csv")
str(data)
summary(data)
```

## Data Cleaning
```{r}
cols_with_zero <- c("Glucose", "BloodPressure", "SkinThickness", "Insulin", "BMI")

data[cols_with_zero] <- lapply(data[cols_with_zero], function(x) {
  ifelse(x == 0, NA, x)
})

data_clean <- na.omit(data)
```

## Exploratory Data Analysis

### Glucose Distribution
```{r}
ggplot(data_clean, aes(x = Glucose)) +
  geom_histogram(binwidth = 10) +
  theme_minimal()
```

### BMI Distribution
```{r}
ggplot(data_clean, aes(x = BMI)) +
  geom_histogram(binwidth = 2) +
  theme_minimal()
```

### Outcome Distribution
```{r}
ggplot(data_clean, aes(x = factor(Outcome))) +
  geom_bar() +
  labs(x = "Diabetes Status", y = "Count") +
  theme_minimal()
```

## Logistic Regression Model
```{r}
model <- glm(Outcome ~ Glucose + BMI + Age + Insulin + BloodPressure,
             data = data_clean,
             family = binomial)

summary(model)
```

## Odds Ratios
```{r}
exp(coef(model))
exp(confint(model))
```

## Prediction
```{r}
probabilities <- predict(model, type = "response")
predicted_class <- ifelse(probabilities >= 0.5, 1, 0)
```

## Model Evaluation
```{r}
conf_matrix <- table(Predicted = predicted_class, Actual = data_clean$Outcome)
conf_matrix
```

## Performance Metrics
```{r}
TN <- conf_matrix[1,1]
FP <- conf_matrix[1,2]
FN <- conf_matrix[2,1]
TP <- conf_matrix[2,2]

accuracy <- (TP + TN) / sum(conf_matrix)
sensitivity <- TP / (TP + FN)
specificity <- TN / (TN + FP)
precision <- TP / (TP + FP)

accuracy
sensitivity
specificity
precision
```

## Confusion Matrix (Caret)
```{r}
confusionMatrix(as.factor(predicted_class), as.factor(data_clean$Outcome))
```

## Conclusion
The model demonstrates that glucose, BMI, and age are the most significant predictors of diabetes. Logistic regression achieved good performance, making it a useful tool for early risk prediction.
